---
title: "Palmer Penguins RSEMINAR"
output:
  html_document: default
  pdf_document: default
date: "2024-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(ggplot2)
```
Palmer Penguins
Properties of the Penguins *living* _in_ Tasmania for the first time.

## Where penguins live

P live on three islands.



```{r islands}
# plot of islands
ggplot(penguins, aes(x = island))+geom_bar(aes(fill=species),position=position_dodge2(preserve="single",reverse= TRUE))
```

Adelie lives on all islands.Interpret. und Auswertung

Question relationship between species and their weight

```{r weight}
ggplot(penguins, aes(x=species, y= body_mass_g))+geom_boxplot()+theme_classic()
````

````{r warning= FALSE, message= FALSE, echo= FALSE}
ggplot(penguins, aes= body_mass_g, y =flipper_length_m)+geom_point(aes(colour=sex))+ geom_smooth(method= "lm")+ theme_classic
````
There is a stro
ng relationship. Yay.
But what is the level of significance of this rel.?

```{r}
fit<- lm(flipper_length_mm ~ body_mass_g + sex, data= penguins)
summary(fit)
````

##welche messungen 
ein odere mehrere messzeitpunkte 
qunat. und keine wdh *lm* mehrere *lmer*
quant.bin (binäre var) keine wdh *glm* mehrere *glmer*
im bsp flipper lenght fl= baseline + einfluus v bodymass (ß) * bodymass selbst +

````{r}
ggplot(penguins, aes (x=body_mass_g, y= flipper_length_mm))+geom_point (aes(colour=sex))+geom_smooth(aes(colour=sex), method ="lm") +scale_color_viridis_d()+theme_classic()
````

sind body masss und flipperlenhght prädiktoren für sex

````{r}
fit <- glm(sex~body_mass_g + flipper_length_mm, family= binomial(), data=penguins)
summary(fit)
````

poission regression anschauen 

of special interest to us is the possible role that the island has on the weight when adjested for the sex and species. This would suggest there would be more food on one island
```{r}
model="lm"(body_mass_g~island + sex + species, data=penguins)
summary (model)

model2=lm(body_mass_g ~sex +species, data= penguins)
anova(model,model2)
```
-14 sagt die auf island dream wiegen 14 g weniger aber nicht significant in t wiegen die 6 gramm mehr als auf b aber auch nihct significant


